\section{Prerequisites}

\begin{defn}[Network]\label{defn_network}
    The network delivers any messages among nodes connected to it, but it may arbitrarily delay or reorder messages.
\end{defn}

\begin{defn}[Well-behaved]
    A node is considered well-behaved if and only if they follow the protocol and is connected to the network.
\end{defn}

\begin{rem}
    A few things that may make it easier to understand FBA and SCP:
    \begin{itemize}
        \item
            $V$ and $v$ are often used as the set of nodes and a node because graph theorists often call nodes vertices.
            Also, $n$ is a common variable name for a natural number, so it is often convenient to use $v$ for a vertex.
        \item
            Pay particular attention to the difference between quorums and quorum slices.
            They sound very similar, and they are related, but they are different.
        \item
            $B$ is often used as a subset of nodes.
            The letter $B$ is used because when we think of a subset, we are often trying to model the set of befouled nodes.
        \item
            $q$ and $Q$ are often used for \textit{quorum slices}, instead of quorums.
            The function $Q$ is for \textit{quorum slices}, not quorums.
            The letter $q$ is often used as a \textit{quorum slice}, and the letter $U$ is often used as a \textit{quorum}.
            I have no idea why that is the case.
        \item
            The letter $x$ is often used as a value, and the letter $a$ is often used as a statement.
    \end{itemize}
\end{rem}
